---
title: "Effect of Car Design on Miles per Gallon"
author: "Stephen J. Osborne"
date: "20th September 2018"
logo: "sjo.png"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, error=FALSE)
```

```{r, echo=FALSE}
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
```

## Introduction

+ Load `mtcars` from datasets

```{r}
library(datasets); data("mtcars")
```

+ The data comprises of fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

+ How does the design of the car affect the miles per gallon (mpg)?

+ I will focus on:
     + Weight, Number of cylinders, Engine displacement, Engine horsepower

## Results: Plots using `ggplotly()`

```{r, echo=FALSE}
d1 <- select(mtcars, mpg, wt, cyl:hp) %>%
     gather(., mpg) %>%
     mutate(., mpgval=rep(mtcars$mpg,4),
            model=rep(rownames(mtcars),4)
            )
colnames(d1) <- c("aspect","value","mpg","model")
d1 <- mutate(d1, aspect=factor(aspect))
levels(d1$aspect) <- c("Cylinders/no units","Displacement/cubic inches","Horsepower/hp","Weight/ton")

plot <- ggplot(d1, aes(x=value,y=mpg)) +
     geom_point(aes(text = paste("Model:", model))) +
     geom_smooth(method="glm",aes(color = aspect)) +
     facet_wrap(~ aspect, scales = "free_x") +
     theme_classic() + theme(legend.position = "none")

ggplotly(plot)
```

## Correlation of Displacement and Power

```{r, echo=FALSE}
d2 <- select(mtcars, mpg, wt, cyl:hp) %>%
     mutate(., model=rownames(mtcars))
fit <- lm(hp ~ disp, data = d2)
plot_ly(d2, x=~disp, y=~hp) %>%
     add_markers(y=~hp) %>%
     add_lines(x=~disp, y=fitted(fit)) %>%
     layout(showlegend = F)
```

## Conclusions

+ All investigated design aspects lower mpg as they increase.
+ Weight is very important as an engine of the same size is required to do more work to move the same mass.
+ Having more cylinders means more fuel is consumed per mile.
+ Displacement and horsepower are positively correlated.
     + Having a greater cylinder volume means the engine is capable of generating more power.
     
